---
title: "Exploring the Relationship between Research and Development (R&D) Expenditure and Economic Growth."
subtitle: ": An Investigation of GNI and GDP Growth with R&D Expenditure in Canadian Data from 2001 to 2023"
author: 
  - Kyungrok Park 
thanks: "Code and data are available at: https://github.com/KyungrokP/Research_budget_and_GNI_relation."
date: today
date-format: long
abstract: "This paper investigates the relationship between Canada's economic growth and R&D expenditure from 2001 to 2023, utilizing data from Statistics Canada. The results of the study indicated that the relationship between Canada's annual R&D expenditure and GDP/GNI growth, as determined through linear regression analysis, was statistically insignificant. Additionally, it was concluded that there exists a negative correlation between nominal share and GDP growth. These insignificant results and negative correlation were explained due to a possible lag effects and issues with R&D budgeting. Additionally, these findings were further justified by referencing the common research results of Kadir, Emir, and Hakan."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(dplyr)
library(here)
library(kableExtra)
library(knitr)
library(modelsummary)
library(stargazer)

```

# Introduction

Since the beginning of the 21st century, various technological
revolutions have occurred in diverse industries worldwide. This has
underscored the importance of investment in scientific technology,
leading not only OECD countries but also nations globally to prioritize
scientific technological investment. A significant aspect of scientific
technological investment is expenditure. Among these, annual R&D budgets
signify a country's dedication to scientific technological investment.
Despite the OECD's average R&D expenditure as a percentage of total GDP
showing a slightly lower percentage of 2.718% in 2022 compared to 2.741%
in 2021 (OECD, 2023), it has exhibited a steady upward trend since 2000.
This indicates that OECD countries are increasingly recognizing the
importance of R&D and steadily increasing their investment.

This study aims to investigate the relationship between R&D expenditure
and GDP or GNI from 2000 to 2022 using Canadian data. It seeks to answer
whether annual R&D spending drives annual GDP, GNI growth, or implies
the country's economic growth for that year in another sense. The
estimand is whether annual R&D expenditure is related to annual GDP or
GNI growth. The data is sourced from Statistics Canada, specifically
quarterly GDP and GNI data for Canada, along with yearly R&D expenditure
information.

While many studies have examined the connection between R&D spending and
economic growth, this paper aims to revisit this correlation, with a
specific focus on distinguishing between GDP and GNI. Furthermore, it
will clearly define the specificity of Canada as a country and examine
the relationship between R&D expenditure and economic growth within the
context of Canada.

This paper consists of four main sections: Data, Model, Results, and
Discussion. In the Data section, we examine the source of the data,how
it was cleaned and utilized, as well as identifying any limitations
inherent in the given dataset. The Method section outlines the
statistical methods, particularly linear regression, utilized to
understand the relationship between annual R&D expenditure and GDP, GNI
growth. It also elucidates the significance and key points of these
statistical methods. The Results section examines the findings regarding
the relationship between Canada's annual R&D expenditure and GDP, GNI
growth, using the statistical methods described in the Model section.
Finally, the Discussion Section critically evaluates the weaknesses of
the analysis employed in this research. Furthermore, it re-evaluates the
results of this study based on the relationship between R&D expenditure
and Economic Growth elucidated in other papers, aiming to confirm the
findings and identify commonalities.

# Data {#sec-data}

As mentioned earlier, both GNI and GDP, along with annual R&D
expenditure, are sourced from Statistics Canada. The reason for
distinguishing between GNI and GDP lies in their definitions.

```{=tex}
\begin{align} 
\text{GDP (Expenditure Approach)} &= C + I + G + (X - M) 
\end{align}
```
```{=tex}
\begin{align} 
\text{GNI} &= \text{GDP} + \text{Income from abroad} 
\end{align}
```
The Expenditure Approach used to calculate GDP, as depicted in the
equation above, sums up a country's total Consumption (C), Investment
(I),Government Spending (G), and Net Exports (X-M). Additionally, GNI
includes income from abroad, added to domestic income, resulting in
GDP + Income from abroad. This means that while GDP growth reflects
overall economic growth within a country, GNI growth also encompasses
the growth of nationals and companies operating abroad. Therefore, while
GDP growth signifies domestic growth, GNI growth reflects the overall
growth and well-being of the nation. Additionally, the relationship
between R&D expenditure and GDP growth will reveal the connection
between R&D expenditure and domestic economic growth, while the
relationship between R&D expenditure and GNI growth will illustrate the
broader implications for overall country development and well-being.

## GDP Growth Data

The GDP data, sourced from Statistics Canada, records quarterly total
expenditure. Through the data cleaning process, quarterly total
expenditures were aggregated into annual totals, enabling the
calculation of GDP growth rates. @tbl-gdp_growth presents the annual GDP
total expenditure and GDP growth rates from 2001 to 2005

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-gdp_growth
#| tbl-cap: "GDP Expenditure and Growth rate of Canada from 2001 to 2005"

GDP_data <- read_csv(file = here::here("data/analysis_data/yearly_GDP_data.csv"))

GDP_data <- mutate(GDP_data, change_in_amount = c(NA, diff(amount_in_mils)))
GDP_data <- mutate(GDP_data, change_rate = ifelse(is.na(change_in_amount), NA, change_in_amount / lag(amount_in_mils) * 100))

GDP_data <- kable(head(GDP_data), col.names = c("Year", "Total Expenditure (million CA $)", "Change in Total Expenditure (million CA $)", "Change rate in total expenditure (%)"))
GDP_data
```

@fig-gdp_growth illustrates Canada's GDP growth from 2001 to 2023. Since
2001, excluding the year 2009 and 2020, Canada's GDP growth has
maintained above 2.8 percent. The year 2020, marked by the severe impact
of the COVID-19 pandemic, recorded a negative growth rate of
approximately -2.67 percent. This stands in stark contrast to the GDP
growth rate of 1.68 percent observed in 2009, following the 2007-2008
financial crisis. Additionally, there was a significant jump in GDP
growth after 2020, and currently, it appears to maintain a growth rate
similar to pre-pandemic levels.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-gdp_growth
#| fig-cap: "Canada's GDP growth Growth rate from 2001 to 2023"

GDP_data <- read_csv(file = here::here("data/analysis_data/yearly_GDP_data.csv"))

GDP_data <- mutate(GDP_data, change_in_amount = c(NA, diff(amount_in_mils)))
GDP_data <- mutate(GDP_data, change_rate = ifelse(is.na(change_in_amount), NA, change_in_amount / lag(amount_in_mils) * 100))
GDP_data <- GDP_data[-1, ]

# Plot the change in rate from 2001 to 2023
library(ggplot2)
ggplot(data = GDP_data, aes(x = year, y = change_rate)) +
  geom_line() +
  labs(x = "Year", y = "GDP growth rate") +
  ggtitle("GDP Growth Rate from 2001 to 2023") +
  theme(plot.title = element_text(hjust = 0.5))
```

## GNI Growth Data

The GNI data, like GDP data, also originates from Statistics Canada. A
crucial aspect here is the method of recording GNI, where Statistics
Canada measured GNI for different years relative to 2017. Specifically,
they set the average GNI for the fourth quarter of 2017 as 100 and
recorded GNI for other years based on this index.

@tbl-gni_growth below presents GNI data from 2001 to 2005. To measure
GNI growth, quarterly GNIs were aggregated by year and compared to the
total GNI of the following year.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-gni_growth
#| tbl-cap: "GNI Growth rate of Canada from 2001 to 2005 based on year 2017 volume"
GNI_data <- read_csv(file = here::here("data/analysis_data/yearly_GNI_data.csv"))

GNI_data <- mutate(GNI_data, change_in_gni = 
                     c(NA, diff(gni_in_2017volume)))

# Calculate the change rate in GNI by year
GNI_data <- mutate(GNI_data, change_rate = 
                     ifelse(is.na(change_in_gni), NA,
                            change_in_gni / lag(gni_in_2017volume) * 100))
GNI_data <- kable(head(GNI_data), col.names = c("Year", "Total GNI (volume index 2017 = 100)", "Change in Total GNI (volume index 2017 = 100)", "Change rate in total GNI (%)"))
GNI_data
```

@fig-gni_growth depicts Canada's GNI growth from 2001 to 2023. Compared
to @fig-gdp_growth, it's noticeable that the volatility of GNI growth is
higher than that of GDP. Particularly, the impact of the 2008-2009
financial crisis is evident in the negative GNI growth of -6.2% in 2009,
as well as the influence of the COVID-19 pandemic reflected in the
-5.56% GNI growth in 2020, showing similar negative growth rates. This
stark contrast with other years' GNI growth rates underscores the
significant impact of these crises. Additionally, there has been
substantial growth in GNI since 2020, indicating concerted efforts to
overcome the negative growth experienced in that year.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-gni_growth
#| fig-cap: "Canada's GNI Growth rate from 2001 to 2023"
GNI_data <- read_csv(file = here::here("data/analysis_data/yearly_GNI_data.csv"))

GNI_data <- mutate(GNI_data, change_in_gni = 
                     c(NA, diff(gni_in_2017volume)))

# Calculate the change rate in GNI by year
GNI_data <- mutate(GNI_data, change_rate = 
                     ifelse(is.na(change_in_gni), NA,
                            change_in_gni / lag(gni_in_2017volume) * 100))

GNI_data <- GNI_data[-1, ]

# Plot the change in rate from 2001 to 2023
library(ggplot2)
ggplot(data = GNI_data, aes(x = year, y = change_rate)) +
  geom_line() +
  labs(x = "Year", y = "GNI growth rate") +
  ggtitle("GNI Growth Rate from 2001 to 2023") +
  theme(plot.title = element_text(hjust = 0.5))
```

## R&D Expenditure Data

Finally, the annual total R&D expenditure data for Canada is sourced
from Statistics Canada. @tbl-rd presents the total R&D expenditure and
change rate from 2001 to 2005, irrespective of the type of investor.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-rd
#| tbl-cap: "Total R&D expenditure and change rate from 2001 to 2005"
research_budget <- read_csv(file = here::here("data/analysis_data/research_budget.csv"))
research_budget <- arrange(research_budget, year)

# Calculate the change in research budget amount by year
research_budget <- mutate(research_budget, change_in_amount = c(NA, diff(amount_in_mils)))

# Calculate the change rate in research budget amount by year
research_budget <- mutate(research_budget, change_rate = ifelse(is.na(change_in_amount), NA, change_in_amount / lag(amount_in_mils) * 100))
research_budget <- kable(head(research_budget), col.names = c("Year", "Total R&D expenditure (million CA $)", "Change in total R&D expenditure (million CA $)", "Change rate in total R&D expenditure (%)"))
research_budget

```

@fig-rd illustrates the change rate in Canada's R&D expenditure from
2001 to 2023. Comparing 2001 to subsequent years, there is a declining
trend in the R&D expenditure change rate between 2002 and 2009. However,
since 2009, aside from a few spikes, there has been a steady increase in
R&D expenditure.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-rd
#| fig-cap: "Canada's Total R&D expenditure change rate from 2001 to 2023"
research_budget <- read_csv(file = here::here("data/analysis_data/research_budget.csv"))
research_budget <- arrange(research_budget, year)

research_budget <- mutate(research_budget, change_in_amount = c(NA, diff(amount_in_mils)))
research_budget <- mutate(research_budget, change_rate = ifelse(is.na(change_in_amount), NA, change_in_amount / lag(amount_in_mils) * 100))

ggplot(data = research_budget, aes(x = year, y = change_rate)) +
  geom_line() +
  labs(x = "Year", y = "Total R&D expenditure change rate ") +
  ggtitle("Total R&D expenditure change rate from 2001 to 2023") +
  theme(plot.title = element_text(hjust = 0.5))

```

## Nominal Share
In addition to the three mentioned data sets above, this study also introduced the nominal share by year. The nominal share refers to the proportion of R&D expenditure relative to the total GDP of the country. If there are years with higher nominal shares, it indicates a greater proportion or contribution to R&D expenditure of the country in that year. @tbl-ns presents the nominal share of Canada from 2000 to 2005. Through these nominal shares, a more precise understanding of Canada's annual contribution to R&D can be obtained, and this will be utilized to assess its correlation with economic growth.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| tbl-cap: Nominal share of Canada between 2000 to 2005
#| label: tbl-ns
research_budget <- read_csv(here("data/analysis_data/research_budget.csv"))
GDP_data <- read_csv(here("data/analysis_data/yearly_GDP_data.csv"))

GDP_data <- mutate(GDP_data, change_in_amount = c(NA, diff(amount_in_mils)))
GDP_data <- mutate(GDP_data, change_rate = ifelse(is.na(change_in_amount), NA, change_in_amount / lag(amount_in_mils) * 100))

research_budget <- mutate(research_budget, change_in_amount = c(NA, diff(amount_in_mils)))
research_budget <- mutate(research_budget, change_rate = ifelse(is.na(change_in_amount), NA, change_in_amount / lag(amount_in_mils) * 100))

nominal_share <- (research_budget$amount_in_mils / GDP_data$amount_in_mils) * 100
GDP_data$nominal_share <- nominal_share
combined_data <- tibble(
  year = GDP_data$year, 
  GDP_amount = GDP_data$amount_in_mils,
  research_budget_amount = research_budget$amount_in_mils,
  nominal_share = nominal_share
)
kable(head(combined_data), col.names = c("Year", "Total GDP (million CA $)", "Total R&D expenditure (million CA $)", "nominal share (%)"))
```
\newpage

# Model

A statistical method used in this study is linear regression. As shown
in the data section, a total of four variables were utilized: GNI
growth, GDP growth, Research budget change rate, and nominal share. In
this study, three linear regression models have been defined to analyze
the relationship between R&D expenditure and economic growth

**First model**: $Y_i\;(dependent\;variable)$ = GNI growth rate (in year
i), $X_i\;(independent\;variable)$= Research budget change rate (in year
i)

**Second model**: $Y_i\;(dependent\;variable)$ = GDP growth rate (in
year i), $X_i\;(independent\;variable)$= Research budget change rate (in
year i)

**Third model**: $Y_i\;(dependent\;variable)$ = GNI growth rate (in year
i), $X_i\;(independent\;variable)$= Nominal share change rate (in year
i)

```{=tex}
\begin{align} 
Y_i = \beta_0 + \beta_1X_i + \varepsilon_i 
\end{align}
```
-   $Y_i$ represents GDP/GNI growth rate in year i

-   $X_i$ represents Research budget/Nominal change rate in year i

-   $\beta_0$ represents the intercept (constant term), the expected
    value of GDP/GNI growth when the independent variables (Research
    Budget rate/Nominal share) is zero.

-   $\beta_i$ represents the coefficient for the R&D expenditure change
    rate and nominal share in year i. It indicates the rate of change in
    GDP/GNI growth rate depending on the change rate in R&D
    expenditure/nominal share

-   $\varepsilon_i$represents the error terms, the discrepancy between
    the actual GDP/GNI growth rate observed and the GDP/GNI growth rate
    predicted by the linear regression model based on the research
    budget/nominal share change rate

\newpage

# Results

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: tbl-lr
#| tbl-cap: "Summary table of 3 different linear regression models"

first_model <- readRDS(here("models/first_model.rds"))
second_model <- readRDS(here("models/second_model.rds"))
third_model <- readRDS(here("models/third_model.rds"))

# Create a list of models
model_list <- list( "GNI growth" = first_model, "GDP growth" = second_model, "GDP growth" = third_model)
modelsummary(model_list, statistic = "(p = {p.value})")
```

@tbl-lr shows the summary of 3 linear regression models. The first model
displays the coefficient of 0.386, which illustrates as R&D expenditure
change rate increases 1%, the GNI growth rate by 0.386 %. The second
model shows the coefficient of 0.088, which also illustrates as R&D
expenditure change rate increases 1%, the GDP growth rate increases by
0.088 %. However, both model shows p-value \> 0.05 which is a generally
accepted value for significance level. Thus, it concludes that the
linear regression result for GNI/GDP growth and R&D expenditure budget
is not significant in this research.

On the other hand, the third model shows coefficient of -12.776 which
shows the negative relationship between GDP growth rate and R&D
expenditure change rate (as nominal share change rate increases by 1 %,
GDP growth rate decreases by -12.776%). The model also shows the p-value
of 0.018 (\<0.05), which suggests the result is significant. However,
this result seems very counter-intuitive since the purpose of R&D
investment is to stimulate the economic growth. Several factors could
influence these results, such as the duration of the data used
(insufficient sample size), lag effects, the background of Canada as a
country, and the purpose and specifics of the R&D investments. Detailed
explanations are provided in the following section, the Discussion
section.

# Discussion

## Lag effects
R&D expenditure is known to have a positive relationship with a country's Total Factor Productivity (TFP) (Zhou&Xia, 2010). However, it is understood that the increase in TFP takes considerable time, leading to long-term growth in a country's GDP/GNI due to the lag effect of TFP. Since this study analyzed the relationship between R&D expenditure and economic growth for each year, it is not immune to the influence of such lag effects. 
In this research, the results of Models 1 and 2 have already indicated that the relationship between R&D expenditure and GNI/GDP growth is not significant. In addition,the negative relationship between nominal share and GDP growth is identified. the results of these three models collectively prove that there is no positive year-by-year relationship between R&D expenditure and economic growth, possibly due to lag effects. Moreover, Canada ranks 6th in the OECD, indicating a high convergence rate. This further underscores that while R&D investment may have a positive effect on economic growth, it requires a considerable amount of time to manifest, likely due to lag effects.

## Possible problems in Resource allocations
The negative relationship between R&D and economic growth revealed in Model 3 of this study may imply issues with the allocation of R&D investment in Canada, beyond just the influence of lag effects. R&D investment involves not only government but also various companies and enterprises. Moreover, the diversity of investor types leads to the allocation of R&D investment across various industries. However, despite this diversity in investment portfolios, changes in R&D allocation may occur annually or quarterly, and concentration in certain industries may negatively impact economic growth.
Conversely, the diversity of investment portfolios may also have a negative impact on economic growth. Canada's productivity declined in 2023, with 10 out of 12 quarters experiencing a decrease (Canadian Chamber of Commerce, 2024). This decline in productivity, coupled with the negative relationship between R&D and economic growth, suggests issues with the allocation of R&D investment in Canada.
Potential solutions include investigating the R&D expenditure and investment levels in each sector to focus investment on industries that can drive the country's growth, or adjusting the annual or quarterly R&D investment portfolio with investment diversity if needed.
To achieve this, proactive intervention in government R&D allocation and advocacy for investment in industries requiring diverse investments may be necessary.


## Relevant Literature
The research, "The Relationship Between Research & Development Expenditures and Economic Growth: The Case of Turkey" (Kadir, Emir & Hakan, 2015), investigated the relationship between R&D expenditure and economic growth using data from Turkey spanning from 1990 to 2013. In detail, Unit root tests, cointegration tests, and Granger’s causality tests were conducted, and the results indicated no significant relationship between R&D expenditure and economic growth. Furthermore, the causality tests revealed no causal relationship betweem them.
In the study by Kadir, Emir & Hakan, the lack of relationship was attributed to the characteristics of Turkey as a country and the lag effect of R&D, which is consistent with the conclusions of this study. So, what does this study imply? The answer can be found in the conclusions mentioned in the previous study. Kadir, Emir & Hakan emphasized the importance of considering both short-term and long-term outcomes of R&D investments. Therefore, R&D investment should be approached with consideration for both quantitative and qualitative standards, and this may necessitate changes in national policies.

## Weaknesses and Next Steps
As mentioned earlier, a weakness of this study lies in exploring the relationship between annual growth and R&D expenditure. Typically, scientific research and development unfold over long periods, making it challenging to predict the accurate timing of the result. For this reason, the relationship between annual growth and scientific investment may not exist, or it might indicate unforeseen directions. Moreover, the realm of R&D investment is vast, encompassing industries not only in general health or labor but also spanning a wide array of scientific fields. However, our study did not differentiate the industries where R&D expenditure was invested but instead aggregated them all to explore their relationship with national growth.
In future research, I would like to delve into segmenting the areas where R&D expenditure is allocated to investigate whether R&D investment in each industry sector contributes to its growth. Furthermore, it is desirable to measure the time it takes for R&D investment to benefit specific industries' growth. This would involve measuring the pace of industrial growth attributable to R&D investment in each sector.


\newpage

# References
---
nocite: '@*'
...
